Autonomous Aerial Exploration for Topological Mapping of Mars Environments

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
edited approach section

Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain. We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig. \ref{fig:transforms}), so $R_c$ can be expressed as
\begin{align*}
R_c=\begin{bmatrix}
\cos\psi & -\sin\psi & 0
\\
\sin\psi & \cos\psi & 0
\\
0 & 0 & 1
\end{bmatrix}, \quad 0\leq\psi<2\pi,
\end{align*}
where $\psi$ represents the angle about the third body-fixed axis. The direction of a 3D measurement may also have a nonzero component in the vertical direction. This is achieved by rotating the sensor frame angle $\theta$ about the second sensor-fixed axis,
\begin{align*}
R_{c,\text{sensor}}=\begin{bmatrix}
\cos\theta & 0 & \sin\theta
\\
0 & 1 & 0
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2},
\end{align*}
where $\Theta$ is the angular range for sensor field-of-view in the vertical direction. Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to Mars as
\begin{align}
z(\psi,\theta)=\begin{bmatrix}\cos\theta\cos\psi & -\sin\psi & \cos\psi\sin\theta
\\
\cos\theta\sin\psi & \cos\psi & \sin\psi\sin\theta
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}
\begin{bmatrix}
\norm{z} \\ 0 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\cos\theta\cos\psi
\\
\cos\theta\sin\psi
\\
-\sin\theta
\end{bmatrix}\norm{z}.
\end{align}
In short, unit vectors are acquired via Euler angles within certain sensor limits.

@@

Next, we show how multiple 3D measurements provide an estimate for information gain. Let $\Psi$ and $\Theta$ be the angular ranges for sensor field-of-view (FOV) in the horizontal and vertical directions, respectively. Within the sensor FOV, we select sample measurements for evaluating their expected information gains. Let the number of sample measurements be $n_\psi$ and $n_\theta$ for horizontal and vertical rotations, respectively. Since $R_c$ and the FOV are known, the set of sample measurement rays, namely $\mathcal{Z}(R_c)$, is known as well. Then, the expected information gain for a single measurement ray $z$ from candidate position $x_c$ and attitude $R_c$ is the negative expected entropy change,
\begin{align}
\label{eqn:Iray}
\mathcal{I}_\text{ray}(x_c,R_c,\psi,\theta)=
\begin{cases}
H(P(r))-\text{E}[H(P(r|x_c,z(\psi+\tilde{\psi},\theta+\tilde{\theta})))], & \mbox{if } z(\psi+\tilde{\psi},\theta+\tilde{\theta})\in\mathcal{Z}(R_c), \\ 
0,                                                                                    & \mbox{otherwise},
\end{cases}
\end{align}
where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}\leq\tilde{\theta}\leq\frac{\pi}{2}$. A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from forward to downward toward the surface of Mars. The expected information gain for the full scan is the summation of expected information gains for individual rays
\begin{align}
\label{eqn:Iscan}
\mathcal{I}_\text{scan}(x_c,R_c)=\sum_{i_\psi=0}^{n_\psi-1}\sum_{i_\theta=0}^{n_\psi-1}\mathcal{I}_\text{ray}\left(x_c,R_c,\frac{\Psi i_\psi}{n_\psi},\frac{\Theta i_\theta}{n_\theta}\right).
\end{align}

@@

\begin{align}
\label{eqn:Proj3DMapComb}
P(\mathbf{m}_{\text{reduced},k})= 
\begin{cases}
    \max_{i\in a_k}{P(\mathbf{m}_i)},	&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh},\\
    \min_{i\in a_k}{P(\mathbf{m}_i)},	& \text{otherwise},
\end{cases}
\end{align}
where initial probability is $0<P_\text{init}<1$ and threshold the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$. This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59}. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.

>>>>

Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain. We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig. \ref{fig:transforms}), so $R_c$ can be expressed as
\begin{align*}
R_c=\begin{bmatrix}
\cos\psi & -\sin\psi & 0
\\
\sin\psi & \cos\psi & 0
\\
0 & 0 & 1
\end{bmatrix}, \quad 0\leq\psi<2\pi,
\end{align*}
where $\psi$ represents the angle about the third body-fixed axis. The direction of a 3D measurement may also have a nonzero component in the vertical direction. This is achieved by rotating the sensor frame angle $\theta$ about the second sensor-fixed axis,
\begin{align*}
R_{c,\text{sensor}}=\begin{bmatrix}
\cos\theta & 0 & \sin\theta
\\
0 & 1 & 0
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2}.
\end{align*}
Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to Mars as
\begin{align}
z(\psi,\theta)=\begin{bmatrix}\cos\theta\cos\psi & -\sin\psi & \cos\psi\sin\theta
\\
\cos\theta\sin\psi & \cos\psi & \sin\psi\sin\theta
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}
\begin{bmatrix}
\norm{z} \\ 0 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\cos\theta\cos\psi
\\
\cos\theta\sin\psi
\\
-\sin\theta
\end{bmatrix}\norm{z}.
\end{align}
In short, unit vectors are acquired via Euler angles within certain sensor limits.

@@

Next, we show how multiple 3D measurements provide an estimate for information gain. Let $\Psi$ and $\Theta$ be the angular ranges for sensor field-of-view (FOV) in the horizontal and vertical directions, respectively. Within the sensor FOV, we select sample measurements for evaluating their expected information gains. Let the number of sample measurements be $n_\psi$ and $n_\theta$ for horizontal and vertical rotations, respectively. Since $R_c$ and the FOV are known, the set of sample measurement rays, namely $\mathcal{Z}(R_c)$, is known as well. Then, the expected information gain for a single measurement ray $z$ from candidate position $x_c$ and attitude $R_c$ is the negative expected entropy change,
\begin{align}
\label{eqn:Iray}
\mathcal{I}_\text{ray}(x_c,R_c,\psi,\theta)=
\begin{cases}
H(P(r))-\text{E}[H(P(r|x_c,z(\psi+\tilde{\psi},\theta+\tilde{\theta})))], & \mbox{if } z(\psi+\tilde{\psi},\theta+\tilde{\theta})\in\mathcal{Z}(R_c), \\ 
0,                                                                                    & \mbox{otherwise},
\end{cases}
\end{align}
where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$. A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from a forward direction to a downward direction toward the surface of Mars. The expected information gain for the full scan is the summation of expected information gains for individual rays
\begin{align}
\label{eqn:Iscan}
\mathcal{I}_\text{scan}(x_c,R_c)=\sum_{i_\psi=0}^{n_\psi-1}\sum_{i_\theta=0}^{n_\psi-1}\mathcal{I}_\text{ray}\left(x_c,R_c,\frac{\Psi i_\psi}{n_\psi},\frac{\Theta i_\theta}{n_\theta}\right).
\end{align}

@@

\begin{align}
\label{eqn:Proj3DMapComb}
P(\mathbf{m}_{\text{reduced},k})= 
\begin{cases}
    \max_{i\in a_k}{P(\mathbf{m}_i)},	&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh},\\
    \min_{i\in a_k}{P(\mathbf{m}_i)},	& \text{otherwise},
\end{cases}
\end{align}
where initial probability is $0<P_\text{init}<1$ and the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$. This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59} for collision avoidance and motion planning. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.
